\textbf{Political Instability and Financial Markets}

### 2017-08-20 doc/paper.tex
Additional updates to sections 1, 2, and 3 (rewriting prose and reframing argument).

We find that financial volatility increases dramatically following (and just before) coup d'etat, assassinations, and resignations. However, these irregular regime changes have disparate effects on the direction of stock returns. Abnormal returns following resignations are large and positive (6\%) while abnormal returns following assassinations are negative and smaller in magnitude (1\%). The impact of coups tends to be negative, but some events such as the temporary ousting of Hugo Chavez during the 2002 Venezuelan coup d'etat attempt result in positive abnormal returns of 20\% or more. This finding is consistent with research that shows that ``good coups'' - coups that lead to democratization and economic liberalization - are not the norm and therefore coups should not be expected to lead to increased economic development and therefore positive market returns \citep{derpanopoulos2015coups}. However, while ``good coups'' may be the minority, they do exist, and coups that attempt to overthrow anti-business autocrats may be expected to result in positive abnormal returns as investors see little risk of a replacement government ``worse'' than the status quo.

Our main finding is therefore that the \textit{type} of political event is less important to the direction of abnormal returns than the \textit{nature} of the event. Events expected to lead to more stable governance, economic liberalization or democratization are associated with positive returns, while those that consolidate authoritarian rule or poor economic policies have the opposite effect.

@@

\singlespacing
\footnotesize
\begin{center}
\begin{longtable}[!ht]{l@{\extracolsep{\fill}}lll}
\caption{List of Coup d'Etat}
\label{tab:coup list}\\
\vspace{-5pt}\\
\hline
\multicolumn{1}{c}{Coup Date}&\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Coup Outcome}\\\hline
\endfirsthead
\multicolumn{3}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\hline
\multicolumn{1}{c}{Coup Date}&\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Coup Outcome}\\\hline
\hline
\endhead
\hline 
\multicolumn{3}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
06/08/1970 & Argentina & Authoritarian to authoritarian\\
03/22/1971 & Argentina & Authoritarian to democratic\\
03/24/1976 & Argentina & Democratic to authoritarian\\
10/06/1976 & Thailand & Democratic to authoritarian\\
10/20/1977 & Thailand & Authoritarian to authoritarian\\
12/12/1979 & South Korea & Authoritarian to authoritarian\\
02/23/1991 & Thailand & Democratic to authoritarian\\
04/05/1992 & Peru & Democratic to authoritarian\\
10/12/1999 & Pakistan & Democratic to authoritarian\\
10/04/2002 & Nepal & Democratic to authoritarian\\
09/19/2006 & Thailand & Democratic to authoritarian\\
01/11/2007 & Bangladesh & Authoritarian to authoritarian\\
04/11/2002 & Venezuela & Failed authoritarian to democratic\\
\hline
\hline
\end{longtable}
\end{center}
\doublespacing
\normalsize

@@

We define the event date as the first trading day in which the market could have reacted to news of the event. For example, during the October 12, 1999 coup d'?tat in Pakistan led by General Pervez Musharraf, the army announced that Prime Minister Nawaz Sharif had been dismissed after market hours at 10:15 pm. We code October 14th, the day in which the market re-opened, as the event day. When events occurred on weekends, we change the event date to the following Monday.

@@

We report abnormal returns separately for coups, assassinations and resignations because they may have distinct effects on stock returns. \autoref{tab:AR-coups} shows abnormal returns for national stock indices both preceding and following coup d'?tat.
Standard errors and p-values are calculated using asymptotic t-statistics as in \citep{mackinlay1997event}. \footnote{It is appropriate to use the standard normal distribution to calculate test statistics because the length of the estimation window is sufficiently long (250 trading days).}

\begin{table}[!ht]
\caption{Abnormal returns following coups} \label{tab:AR-coups}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-coups-car.txt}
  \hline
\ExpandableInput{../tables/artable-coups-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative).
\end{threeparttable}
\end{center}
\end{table}

Event day ARs are negative for all coups except the 1971 and 1976 coups in Argentina. The 1976 coup in Argentina appears to be somewhat of an anomaly.  The ARs on the event day are larger than for any other regime change in the sample (28\%). It is also the only coup with statistically significant (and extremely large) 7-day and 20-day pre-event CARs. Finally, although the event day abnormal return is large and positive, the 7-day and 20-day post-event CARs are negative.

Event day ARs for the 1970 coup in Argentina, the 1979 coup in South Korea, the 1991 coup in Thailand, the 1992 coup in Peru, and the 1999 coup in Pakistan are all negative and statistically different than zero. Moreover, all of these cases except Thailand have negative post-event CARs and pre-event CARs that are statistically indistinguishable from zero.

If the 1976 coup in Argentina is dropped from the sample, the average coup has a $-2.7\%$ event day AR. Even so, markets usually recovered quickly: the market returned to its pre-coup level following a negative event day stock return within 5 trading days in all but one case.

Although it is tempting to dismiss the 1976 Argentina coup as an outlier, there is reason to to believe that coups can in fact result in large positive abnormal returns. The 2002 failed coup d'?tat attempt against Hugo Chavez is a case in point. On the evening of April 11, 2002, coup plotters removed Chavez from office and later detained him. Pedro Carmona, a Venezuelan business leader, was named the transitional President of Venezuela. Two days later, on April 13, 2002, a popular uprising led to Chavez's reinstatement as president.

The event is a perfect natural experiment because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated. It provides an estimate of the market's valuation of a transition from the Chavez regime to the Carmona regime and its valuation of a transition from the Carmona regime back to the Chavez regime. It is therefore a good example of the impact of a shift from a left-wing populist government to a pro business regime on the value of the largest firms in an emerging market.

@@

\begin{figure}[!ht]
\includegraphics{../figs/venezuela_coup_attempt_2002.pdf}
\caption{Abnormal returns surrounding the 2002 Venezuelan coup d'?tat attempt}
\label{fig:AR-Ven}
\end{figure}

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those table \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like coups, there is some evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven almost entirely by three events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

\begin{table}[!ht]
\caption{Abnormal returns following assassinations} \label{tab:AR-ass}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-ass-car.txt}
  \hline
\ExpandableInput{../tables/artable-ass-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative).
\end{threeparttable}
\end{center}
\end{table}

There is no evidence of post or pre-event CARs in almost any of the assassinations. As with coups, the number of days that it took the stock market to rebound to pre-event levels is fairly low. One exception is the assassination of William Mckinley in which the stock market didn't fully recover for 963 days, or almost 4 calendar years. However, this was likely caused by the Panic of 1901, which began when the stock market crashed on May 17th, 1901, and not by McKinley's death (although the assassination may have exacerbated the panic). In any case, the length of this time period is so long that we omitted it when calculating the mean days to rebound in the figure.

As shown in \autoref{tab:AR-resignations}, and in contrast to coups and assassinations, abnormal returns following resignations are large and positive. The mean event day abnormal return is a little over 6\% and event day abnormal returns are only negative and statistically significant at even the ten percent level in one out of the fifteen resignations (Tunisia on June 31, 2011). But positive stock returns are short lived: the 7 day cumulative abnormal return is nearly identical to the event day abnormal return and the 20 day cumulative abnormal return is not statistically different from zero.

\begin{table}[!ht]
\caption{Abnormal returns following resignations} \label{tab:AR-resignations}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-res-car.txt}
  \hline
\ExpandableInput{../tables/artable-res-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative).
\end{threeparttable}
\end{center}
\end{table}

The positive event day abnormal return following resignations is not surprising since resignations occurred because of ``poor performance and/or loss of authority.'' Leaders were often ousted following corruption charges, financial crises, and/or political violence. For example, consider Suharto's resignation from office as the President of Indonesia on May 1998, just after the 1997 Asian Financial Crisis.

Prior to the crisis, the Indonesian economy suffered from a number of ailments including crony capitalism, poor corporate governance, growing short-term debt and a corrupt banking system \citep{radelet2000onset}. In addition, Suharto's ill health created considerable policy uncertainty, which only added to the financial panic in Indonesia. Indonesia was also hesitant to adopt structural reforms which further weakened its economic image. As a result, the crisis gave additional ammo to opposition leaders and public opposition culminated in riots in mid-May 1998. Suharto eventually lost his military support and was forced to resign.

\subsection{Volatility} \label{subsec: Volatility}
Although irregular regime changes have disparate effects on the direction of stock returns, the results from \autoref{subsec: Abnormal Returns} suggest that all of these political events increase financial volatility. However, since stock volatility is not directly observable, one must decide how to best estimate volatility. My estimates are obtained from a generalized autoregressive conditional heteroskedasticity (GARCH) model estimated using 1000 pre-event days, the event day and 1000 post-event days. As in \citep{jensen2005market} and \citep{leblang2005government}, we use the GARCH (1,1) specification. In particular, for national stock index $i$,
\begin{align*}
R_{it}=\mu_i + \epsilon_{it},\hspace{1cm} \epsilon_{it}\sim \mathcal{N}\left(0,\sigma_{it}^2\right),
\end{align*}
where $\mu_i$ is a constant and,
\begin{align*}
\sigma_{it}^2&=\gamma_{i}+\alpha_{i}\epsilon_{i,t-1}^2+\beta_{i}\sigma_{i,t-1}^2.
\end{align*}
The key parameter of interest is the conditional variance, $\sigma_{it}^2$. The one-period-ahead volatility forecasts, $\sigma_{it}$, are larger when $\epsilon_{i,t-1}^2$ and $\sigma_{i,t-1}^2$ are larger. In other words, the model predicts that large shocks will be followed by other large shocks.

@@

The third concern, which is that the political events in this paper are not unexpected, is addressed in a number of ways. First, one can note that neither the 7-day or the 20-day mean CAR from tables \autoref{tab:AR-coups}, \autoref{tab:AR-ass} or \autoref{tab:AR-resignations} are statistically different than zero at even the 10\% level. Although some of the irregular regime changes have statistically significant pre-event CAR's, most of them do not, and the ones that do do not always move in the same direction. For example, there was a negative 11.8\% 7-day CAR preceding the resignation of President Estrada of the Philippines in January of 2001 and a positive 14.7\% 7-day CAR preceding the resignation of President Fernando de la R?a of Argentina in December of 2001, even though both resignations were associated with large positive event day abnormal returns.

@@

While the Egyptian Revolution may have positive pro-democracy effects on the Egpytian political system, its short-term impact on the economy was disastrous. As shown in \autoref{fig:CAR-Egypt}, abnormal returns the Egyptian Stock Exchange Index (EGX 30) were around -7\% on January 26th and -10\% the day after. To prevent further decline during the uprising, the Egyptian Stock Exchanged closed at the end of trading on January 27th. President Mubarak resigned on February 11, but the market remained closed until March 23, when CARs declined by another 9\%, before rebounding slightly thereafter.

@@

An important question is whether other popular uprisings have had similar adverse economic consequences. To examine this, we explore all resignations that were driven by significant public protests.\footnote{The set of resignations includes all those listed in either the Coup d'?tat Events Handbook or the Archigos Version 2.9 data set with available financial data. In practice, this is the 2011 Egyptian Revolution and the list or resignations in \autoref{tab:AR-resignations}.} Public protests include popular demonstrations, riots, non-violent civil resistance and other forms of public discontent. These events are listed in \autoref{tab:protest-list} in the appendix.

>>>>

We find that financial volatility increases dramatically following (and just before) coup d'etat, assassinations, and resignations. However, unlike the previously mentioned studies, we find that these irregular regime changes have disparate effects on the direction of stock returns. Abnormal returns following resignations are large and positive (6\%) while abnormal returns following assassinations are negative and smaller in magnitude (1\%). The impact of coups tends to be negative, but some events result in positive abnormal returns of 20\% or more. This finding is consistent with research that shows that ``good coups'' - coups that lead to democratization and economic liberalization - are not the norm and therefore on average coups should not be expected to lead to increased economic development and positive market returns \citep{derpanopoulos2015coups}. However, while ``good coups'' may be the minority, they do exist, and coups that attempt to overthrow anti-business autocrats may be expected to result in positive abnormal returns as investors see little risk of a replacement government ``worse'' than the status quo.

Our main finding is therefore that the type of political event and its expected impact on economic policy determines the direction of abnormal returns. Events expected to lead to more stable governance, economic liberalization, or democratization (such as willful resignations and coups that overthrow protectionist or leftist autocrats) are associated with positive returns, while those that consolidate authoritarian rule, exacerbate poor economic policies, or merely increase policy uncertainty have the opposite effect.

@@

\singlespacing
\footnotesize
\begin{center}
\begin{longtable}[!ht]{l@{\extracolsep{\fill}}lll}
\caption{List of Coup d'Etat}
\label{tab:coup list}\\
\vspace{-5pt}\\
\hline
\multicolumn{1}{c}{Coup Date}&\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Coup Outcome}\\\hline
\endfirsthead
\multicolumn{3}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\hline
\multicolumn{1}{c}{Coup Date}&\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Coup Outcome}\\\hline
\hline
\endhead
\hline 
\multicolumn{3}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
06/08/1970 & Argentina & Authoritarian to authoritarian\\
03/22/1971 & Argentina & Authoritarian to authoritarian\\
03/24/1976 & Argentina & Democratic to authoritarian\\
10/06/1976 & Thailand & Democratic to authoritarian\\
10/20/1977 & Thailand & Authoritarian to authoritarian\\
12/12/1979 & South Korea & Authoritarian to authoritarian\\
02/23/1991 & Thailand & Democratic to authoritarian\\
04/05/1992 & Peru & Democratic to authoritarian\\
10/12/1999 & Pakistan & Democratic to authoritarian\\
10/04/2002 & Nepal & Democratic to authoritarian\\
09/19/2006 & Thailand & Democratic to authoritarian\\
01/11/2007 & Bangladesh & Authoritarian to authoritarian\\
04/11/2002 & Venezuela & Failed authoritarian to democratic\\
\hline
\hline
\end{longtable}
\end{center}
\doublespacing
\normalsize

@@

We define the event date as the first trading day in which the market could have reacted to news of the event. For example, during the October 12, 1999 coup d'etat in Pakistan led by General Pervez Musharraf, the army announced that Prime Minister Nawaz Sharif had been dismissed after market hours at 10:15 pm. We code October 14th, the day in which the market re-opened, as the event day. When events occurred on weekends, we change the event date to the following Monday.

@@

We report abnormal returns separately for coups, assassinations and resignations because they may have distinct effects on stock returns. \autoref{tab:AR-coups} shows abnormal returns for national stock indices both preceding and following coup d'etat.
Standard errors and p-values are calculated using asymptotic t-statistics as in \citep{mackinlay1997event}. \footnote{It is appropriate to use the standard normal distribution to calculate test statistics because the length of the estimation window is sufficiently long (250 trading days).}

\begin{table}[!ht]
\caption{Abnormal returns following coups} \label{tab:AR-coups}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-coups-car.txt}
  \hline
\ExpandableInput{../tables/artable-coups-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative). Returns are inflation adjusted. 
\end{threeparttable}
\end{center}
\end{table}

The average coup has a $-2.5\%$ event day AR. Event day ARs for the 1970 coup in Argentina, the 1979 coup in South Korea, the 1991 coup in Thailand, the 1992 coup in Peru, and the 1999 coup in Pakistan are all negative and statistically different than zero. Moreover, all of these cases except Thailand have negative post-event CARs and pre-event CARs that are statistically indistinguishable from zero. In all of these cases, the coup in question either overthrew a democratically elected government or changed governance from one military ruler to another. The initial negative reaction followed by additional post-event negativity is consistent with the expected market reaction from a successful authoritarian coup followed by post-event consolidation of power.

Event day ARs are negative for all coups except the 1971 coup in Argentina and the 2002 failed coup d'etat attempt against Hugo Chavez. These results provide evidence that that coups can in fact result in positive abnormal returns. While the 1971 Argentinian coup did result in another military leader, it did so while calling for free and democratic elections and replaced a government that had adopted extreme protectionist economic policies. The ultimately failed coup against Hugo Chavez replaced a left-wing populist government with a new pro-business president.

The failed coup attempt against Hugo Chavez is a perfect natural experiment because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated. On the evening of April 11, 2002, coup plotters removed Chavez from office and later detained him. Pedro Carmona, a Venezuelan economist and business leader, was named the transitional President of Venezuela. Two days later, on April 13, 2002, a popular uprising led to Chavez's reinstatement as president. This provides an estimate of the market's valuation of a transition from the Chavez regime to the Carmona regime and its valuation of a transition from the Carmona regime back to the Chavez regime. By extension, it provides an estimate of the impact of a shift from a left-wing populist government to a pro business regime in an emerging market.

@@

\begin{figure}[!ht]
\includegraphics{../figs/venezuela_coup_attempt_2002.pdf}
\caption{Abnormal returns surrounding the 2002 Venezuelan coup d'etat attempt}
\label{fig:AR-Ven}
\end{figure}

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those table \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven almost entirely by three events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

These results are consistent with our hypothesis that the nature of the political event and its expect impact on policy matters. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination is typically a random event that will not necessarily cause immediate change in economic policy. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

\begin{table}[!ht]
\caption{Abnormal returns following assassinations} \label{tab:AR-ass}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-ass-car.txt}
  \hline
\ExpandableInput{../tables/artable-ass-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative). Returns are inflation adjusted. 
\end{threeparttable}
\end{center}
\end{table}

There is no evidence of post or pre-event CARs in almost any of the assassinations. This is consistent with expectations as assassinations are typically not predictable. As with coups, the number of days that it took the stock market to rebound to pre-event levels is fairly low.\footnote{One exception is the assassination of William Mckinley in which the stock market didn't fully recover for 963 days, or almost 4 calendar years. However, this was likely caused by the Panic of 1901, which began when the stock market crashed on May 17th, 1901, and not by McKinley's death (although the assassination may have exacerbated the panic). In any case, the length of this time period is so long that we omitted it when calculating the mean days to rebound in the figure.}

In contrast to coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{tab:AR-resignations}). The mean event day abnormal return is over 6\% and event day abnormal returns are only negative and statistically significant at even the ten percent level in one out of the fifteen resignations (Tunisia on June 31, 2011). However, these positive returns are short lived: the 7 day cumulative abnormal return is nearly identical to the event day abnormal return and the 20 day cumulative abnormal return is not statistically different from zero.

\begin{table}[!ht]
\caption{Abnormal returns following resignations} \label{tab:AR-resignations}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-res-car.txt}
  \hline
\ExpandableInput{../tables/artable-res-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative). Returns are inflation adjusted. 
\end{threeparttable}
\end{center}
\end{table}

These results are also consistent with the idea that the expected impact of a political event on economic policy is highly important to the directionality of returns. The positive event day abnormal return following resignations is not surprising since resignations occurred because of ``poor performance and/or loss of authority.'' Leaders were often ousted following corruption charges, financial crises, and/or political violence. For example, consider Suharto's resignation from office as the President of Indonesia on May 1998, just after the 1997 Asian Financial Crisis.

Prior to the crisis, the Indonesian economy suffered from a number of ailments including crony capitalism, poor corporate governance, growing short-term debt and a corrupt banking system \citep{radelet2000onset}. In addition, Suharto's ill health created considerable policy uncertainty, which only added to the financial panic in Indonesia. Indonesia was also hesitant to adopt structural reforms which further weakened its economic image. As a result, the crisis gave additional ammunition to opposition leaders, and public opposition culminated in riots in mid-May 1998. Suharto eventually lost his military support and was forced to resign.

\subsection{Volatility} \label{subsec: Volatility}
Although irregular regime changes have disparate effects on the direction of stock returns depending on expected policy outcomes, the results from \autoref{subsec: Abnormal Returns} suggest that all of these political events increase financial volatility. However, since stock volatility is not directly observable, one must decide how to best estimate volatility. Our estimates are obtained from a generalized autoregressive conditional heteroskedasticity (GARCH) model estimated using 1000 pre-event days, the event day and 1000 post-event days. As in \citep{jensen2005market} and \citep{leblang2005government}, we use the GARCH (1,1) specification. In particular, for national stock index $i$,
\begin{align*}
R_{it}=\mu_i + \epsilon_{it},\hspace{1cm} \epsilon_{it}\sim \mathcal{N}\left(0,\sigma_{it}^2\right),
\end{align*}
where $\mu_i$ is a constant and,
\begin{align*}
\sigma_{it}^2&=\gamma_{i}+\alpha_{i}\epsilon_{i,t-1}^2+\beta_{i}\sigma_{i,t-1}^2.
\end{align*}
The key parameter of interest is the conditional variance, $\sigma_{it}^2$. The one-period-ahead volatility forecasts, $\sigma_{it}$, are larger when $\epsilon_{i,t-1}^2$ and $\sigma_{i,t-1}^2$ are larger. In other words, the model predicts that large shocks will be followed by other large shocks.

@@

The third concern, which is that the political events in this paper are not unexpected, is addressed in a number of ways. First, one can note that neither the 7-day or the 20-day mean CAR from tables \autoref{tab:AR-coups}, \autoref{tab:AR-ass} or \autoref{tab:AR-resignations} are statistically different than zero at even the 10\% level. Although some of the irregular regime changes have statistically significant pre-event CAR's, most of them do not, and the ones that do not always move in the same direction. For example, there was a negative 11.8\% 7-day CAR preceding the resignation of President Estrada of the Philippines in January of 2001 and a positive 14.7\% 7-day CAR preceding the resignation of President Fernando de la Rua of Argentina in December of 2001, even though both resignations were associated with large positive event day abnormal returns.

@@

While the Egyptian Revolution may have positive pro-democracy effects on the Egpytian political system, its short-term impact on the economy was disastrous. As shown in \autoref{fig:CAR-Egypt}, abnormal returns on the Egyptian Stock Exchange Index (EGX 30) were around -7\% on January 26th and -10\% the day after. To prevent further decline during the uprising, the Egyptian Stock Exchanged closed at the end of trading on January 27th. President Mubarak resigned on February 11, but the market remained closed until March 23, when CARs declined by another 9\%, before rebounding slightly thereafter.

@@

An important question is whether other popular uprisings have had similar adverse economic consequences. To examine this, we explore all resignations that were driven by significant public protests.\footnote{The set of resignations includes all those listed in either the Coup d'etat Events Handbook or the Archigos Version 2.9 data set with available financial data. In practice, this is the 2011 Egyptian Revolution and the list or resignations in \autoref{tab:AR-resignations}.} Public protests include popular demonstrations, riots, non-violent civil resistance and other forms of public discontent. These events are listed in \autoref{tab:protest-list} in the appendix.
